---
title: "Stress Detection in Wearables"
author: "Alice Liao, Anna Darwish, Bernardo Martinez"
date: "9/19/2020"
output: 
  pdf_document:
     latex_engine: xelatex
     number_sections: true
---

```{r, function checking for installed packages, echo=FALSE, results="hide", include=FALSE}
# Validate that all necessary packaged have been downloaded, install otherwise or throw err package DNE
pkgTest <- function(x)
{
  if (!require(x,character.only = TRUE))
  {
    install.packages(x,repos = "http://cran.r-project.org", dep=TRUE)
    if(!require(x,character.only = TRUE)) stop("Package not found")
  }
}
```

```{r Package Test Libraries, echo=FALSE, results="hide", include=FALSE}
pkgTest("dplyr")
pkgTest("ggplot2")
pkgTest("tidyr")
pkgTest("knitr")
pkgTest("kableExtra")
```

```{r Load in Libraries, echo=FALSE, results="hide", include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
ggplot2::theme_set(new = theme_bw())
```

```{r load data, include=F}
wesad <- read.csv("merged_wesad_data.csv")

wesad <- wesad %>%
  mutate(label = as.factor(label),
         subject = as.factor(subject))

head(wesad)
```

# Introduction

In the past few years, the use of wearables has exploded [[1]][Bibliography]. With it, researchers around the globe are exploring how tracking health information via wearables, such as a person's heart rate, can help gain insight into their physical and emotional states [[2]][Bibliography]. Across many different functionalities, such as tracking workouts or monitoring sleep, wearables also present an exciting opportunity to detect stress. In doing so, wearable users today, such as Apple watch user Jason Hiner, have become more cognizant of their emotional state and worked to improve it [[3]][Bibliography]. Given that stress can lead to long term health issues, such as heart disease, high blood pressure, diabetes, and depression or anxiety [[4]][Bibliography], it is critical for us to ensure that wearables can accurately detect stress and distinguish it from other affective states.

However, researchers have found evidence to be skeptical of the accuracy and insights a wearable watch is capable of offering [[5]][Bibliography]. For example, experts at the Cleveland Clinic found that heart-rate data in many popular wearables can be wrong 10-20% of the time [[6]][Bibliography]. While research has been done to examine whether these sensors are capable of detecting excited affective states, such as stress, relative to a subject's baseline state, there are far fewer studies that examine whether these sensors are capable of distinguishing excited affective states from one another, such as stress vs amusement [[7]][Bibliography]. A recent paper by Schmidt et. al was able to generate highly accurate machine learning models that classified wearers' states among baseline, stress, and amusement conditions using an Empatica 4 wristband [[7]][Bibliography]. However, these methods were unable to provide significant insight into the relationship between the most informative metrics and these affective states.

In this paper we hope to expand upon the work done within the Schmidt et. al paper by delving deeper into the relationship between the metrics collected by these sensors and the wearers' affective states with a logistic regression model. Specifically, we hope to accomplish the following goals:

1. Determine whether sensor data are useful in discriminating between stress and amusement conditions
2. Understand and describe how sensor data are useful in discriminating between stress and amusement
3. Determine which types and combination(s) of sensor data are most useful in discriminating between stress and amusement
4. Determine whether we can detect stress (vs amusement) only using the wrist-worn wearable
5. Quantify the heterogeneity across individuals in the response to stress versus amusement

# Data

## Collection Overview

For the purposes of this paper, we will use the biological definition of stress: an affective state defined to be the 'nonspecific response of the body to any demand upon it' [[7][Bibliography]. The data used in this paper was collected by attaching a *RespiBAN Professional* chest device and *Empatica E4* wristband to 17 graduate students. Due to sensor malfunctions, two of the subjects' data had to be discarded, for a final total of 15 subjects. Three of these subjects were women while the rest were men, and their ages ranged from 25 years old to 35 years old. This experiment sought to elicit three different affective states: neutral, stress, and amusement, and accomplished this by having the subjects perform stressful tasks (e.g public speaking) and amusing tasks (e.g watching funny videos) with breaks of meditation between sessions to restore subjects to their baseline. The stress-condition tasks and amusement-condition tasks were interchanged to avoid effects of order [[7]][Bibliography].

In the Appendix, you can see a full data dictionary in Table 1 that defines the metrics, their sampling rates, and any extracted features. Finally, it is important to note that for Subject 3, there appears to have been a brief malfunction (over a span of approximately 0.1 seconds during stressed condition) that led to impossible measurements, such as temperatures of -273$^\circ$ C. We chose to treat these ranges as missing values and also used linear interpolation to replace these faulty measurements.

## Exploratory Data Analysis

We plotted the distributions of predictors for each of the subjects and realized for some people, certain metrics serve as perfect separators for amused vs. stressed conditions (i.e. the distributions under the two conditions are significantly different or separated), but that's not always the case. Among all predictors, electrodermal activity (EDA) and temperature are the two physiological signals that appear to be separators for most of the subjects. 

```{r loadEDAgraphs, fig.width=10, fig.height=4, message=F, echo=F}
# distribution of average body temperature measured at chest (window = 60s, shift = 0.25s)
ggplot(wesad) +
  geom_histogram(aes(x=Temp_CHEST_MEAN, fill=label),alpha = 0.6, position = "identity")+
  facet_wrap(~subject, nrow = 3, scales = "free")+
  labs(x = "Body Temperature (??C)",
       title = "Figure 1: RespiBan Skin Temperature") 
# distribution of average electrodermal activity measured at wrist (window = 60s, shift = 0.25s)
ggplot(wesad) +
  geom_histogram(aes(x=EDA_CHEST_MEAN, fill=label),alpha = 0.6, position = "identity")+
  facet_wrap(~subject, nrow = 3, scales = "free")+
  labs(x = expression(paste("Electrodermal Activity (",mu,"S)")),
       title = "Figure 2: RespiBan Electrodermal Activity")
```

From Figure 1 we see body temperatures are different in amused vs stressed states. For some subjects, such as Subject 10 and 11, the difference is more pronounced in that the distributions of body temperatures are completely separated by the two states. Similarly in Figure 2 we observe different electrodermal acitivities in different emotional states. The histograms also suggest not everyone reacts to amusement and stress in the same way. In general, it seems more sweat (higher EDA) is associated with stress, but Subject 14 could have high EDA but was amaused too. As for temperature, some subjects had higher temperature under stress (e.g. Subject 10 and Subject 15) whereas some had lower temperature when amused (e.g. Subject 11 and Subject 17). Skin temperature and EDA measured at wrist appear to be separators too and their distributions can be found in Appendix.  

# Methods

Our first step with this dataset involved aligning the *RespiBAN* data and *Empatica E4* data. For the *RespiBAN* data, each metric was collected at 700Hz. Across all metrics, 700Hz was well above the minimum sampling rate [[8]][Bibliography], [[9]][Bibliography], [[10]][Bibliography], [[11]][Bibliography], which is defined to be the minimum number of samples needed to glean precise feature extractions. For more detail on the minimum sampling rates (MSRs) across the different metrics, please reference Table 1 in the Appendix. Despite these different MSRs, these metrics ultimately needed to be aligned to represent distinct and whole samples. We selected to downsample to 4 Hz, which has been documented as the minimum sampling rate of electrodermal activity (EDA). From our data exploration, EDA appeared to be a key metric in distinguishing stress and amusement. While there is controversy on whether higher sampling rates are required for electrodermal activity [[!!!! FIND REF !!!]], our exploratory data analysis showed that there were stark differences in the electrodermal activity between the two states for many of the subjects. We also chose this value, as temperature, which has an MSR of 1 Hz, also appeared to be a critical indicator. 

We accomplished this downsampling by calculating our features over a window of one minute with a shift of 0.25 seconds. This was motivated by the Schmidt et. al's paper's window and shift decisions. Please refer to Table 1 in the Appendix for more details on other transformations we explored in generating our models. (REMOVE IF ULTIMATELY FALSE) Please reference Table ** later on to compare the results of our final models to those with a sampling rate of 2 Hz for sensitivity analysis, which we accomplished with a 60 second window and 0.5 second shift.

For this paper, we have decided to generate several logistic regression models to better understand and interpret the relationship between these metrics. We began by exploring histograms and scatterplots of our metrics. From this we started with a model with the following predictors: RespiBAN EDA, RespiBAN Body Temperature, and Subject and ran 10-fold cross-validation on our dataset to estimate the test error. Note that it was necessary to include the subject id in order for us to be able to treat our observations as independent. Then we performed forward model selection to check if we missed any other important predictors that can improve the accuracy of prediction. In addition, since both *Empatica E4* wrist sensor and *RespiBan* chest sensor provide readings for EDA and Body Temperature, we can compare which type of sensor is more useful in discriminating between stress and amusement and if wrist data alone is enough. 

# Results

With the model final model using the predictors outlined previously we were able to achieve an accuracy of 0.8995 when using a 0.5 threshold to categorize the different probabilities outputted by the model.  The coeffcients interpretations for the non-subject variables are as follows: 

1.) For every one unit increase in EDA measurement , the log-odds of the subject being stressed increases by 24.75626
2.) For every one unit increase in the temperature taken from the chest of the subject, the log-dds of the subject being stressed decreases by 1.14694.
3.) For every one increase in one of the variables while holding the other constant, the log-odds of the subject being stressed decreases by 0.50790.

This model also provided statistically significant values for all the coefficients used. 

```{r loadmodelandplots}
model.interaction = readRDS("Model.RDS")
threshold <- 0.5
model.interaction <- augment(model.interaction, type.predict = "response")

model.interaction %>%
  mutate(stressed_predict = if_else(.fitted > threshold, 1, 0)) %>%
  group_by(label, stressed_predict) %>%
  summarise(n = n()) %>%
  spread(label, n) %>%
  kable(format="markdown")
```

Here we see a confusion matrix of the predicitions of our model [ADD MORE DETAILS WHEN FINAL MODEL IS DONE]

Below we see the ROC curve which illustrates the diagnostic ability of the model as the discrimination threshold is varied. This plot also allows us to determine an AUC value which is another measure of performance of the model. [ADD DETAILS OF WHAT SPECIFICALLY IS SHOWN IN PLOTS WHEN FINAL MODEL IS DETERMINED]

```{r}
ggplot(model.interaction, aes(d = as.numeric(label), m = .fitted)) + 
  geom_roc(n.cuts = 0) + 
  geom_abline(intercept = 0)

roc(model.interaction$label,model.interaction$.fitted)$auc
```

Lastly we look at the binned residual plots to as a means of model diagnostics. We are specifically going to look for patterns, assess any possible non-linear trends, and look at the overall magnitude of the residuals in each of the bins. 

```{r}
arm::binnedplot(x=model.interaction$.fitted,y=model.interaction$.resid,
                xlab="Predicted Probabilities", 
                main = "Binned Residual vs. Predicted Probabilities")

arm::binnedplot(x=model.interaction$EDA_CHEST_MEAN,y=model.interaction$.resid,
                xlab="RespiBAN Electrodermal Activity", 
                main = "Binned Residual vs. RespiBAN Electrodermal Activity")

arm::binnedplot(x=model.interaction$Temp_CHEST_MEAN,y=model.interaction$.resid,
                xlab="RespiBAN Body Temperature", 
                main = "Binned Residuals vs. RespiBAN Body Temperature")
```

# Discussion

Based on the model created previously and its very high accuracy we can confidently say that sensor data is useful in discriminating between the different conditions. There are some limitations to this answer mainly in the form the data is collect [ELABORATE MORE ON DATA LIMITATIONS HERE]. 
  
It is important to note that not all of the sensor data is useful in discriminating between the two states as shown previosuly in the EDA portion of the report. To validate this claim we used a forward selection method on the previous model in order to see if there was any avenue for bettering it by adding more sensor variables. This method used AIC as a metric from which to compare the models generated and it did not change our model. That leads us to conclude that the sensor data that is most useful in determining between stress and amusement are the body temperature of the individual and electrodermal activity taken from the chest. [CHANGE TO SIGNIFICANT PREDICTORS FROM FINAL MODEL]

Given that there are two types of sensors we also find it interesting to understand whether it not it is necessary to have both sensors to determine. Apart from the model mentioned previously, we also built a model that just takes into consideration the wrist data. Using insights from the EDA portion of the report along with using a forward selection method to determine the best model using AIC as a metric, led to a model that uses electrodermal activity and temperature taken from the wrist. This model overall led to worse results than the model using chest data with an accuracy of [INSERT WRIST MODEL ACCURACY HERE] and an AUC of [INSERT WRIST MODEL AUC HERE. The resulting accuracy may be of use especially in more casual settings but it is not as accurate and may not be of use where the prediction has serious implications. Some possibilities for this can stem from the accuracy of the sensors on the wrist wearable. For example, it is much easier to take on and off the wrist wearable along with it being more subject to movements and other factors that effect the sensors. For more detailed plots regarding confusion matrices, ROC curves, and binned residual plots for this model can be found in the appendix.

Lastly, an important question to answer is the heterogeneity across individuals in the response to stress versus amusement. All humans are different, and reactions to stress and amusement produce different physical reactions. One way to quantify this question is by analyzing the results of our model that uses the specific subject as another variable. By analyzing the different coefficients we saw that many of these coefficients were different by a significant margin. For example, [SHOW EXAMPLE OF MAJOR DIFFERENCE IN COEFFICIENT]. Another we tried answering this question was exploring the different predictors that the model uses and checking how their plots differ across subjects. Our results show that [INSERT INSIGHTS FROM ]

1. Determine whether sensor data are useful in discriminating between stress and amusement conditions
2. Understand and describe how sensor data are useful in discriminating between stress and amusement
3. Determine which types and combination(s) of sensor data are most useful in discriminating between stress and amusement
4. Determine whether we can detect stress (vs amusement) only using the wrist-worn wearable
5. Quantify the heterogeneity across individuals in the response to stress versus amusement

# Conclusion

To summarize our final results, we found that ... were the most useful predictors in distinguishing stress from amusement in these subjects. Further, we found that the data from the *Empatica E4* wrist sensor alone was ____ than the data from the *Empatica E4* and *RespiBAN* combined. Finally, for the predictors ..., there was a significant interaction effect between that metric and at least one of the subjects. This suggests that there may be other covariates or metrics that are required to make a generalizable model.

Several of the limitations in the analysis include our inability to cater to each metric's preferred sampling rate, the limited number of affective states, the imprecision in the devices themselves, and the lack of representation from younger and older age groups, and the limited representation of women. For future studies, it will be necessary to examine whether sensor data from wearables is capable of distinguishing other affective states that yield similar physiological responses, such as anger, and to include subjects from a broad range of ethnicities, races, and age groups. Another possible avenue to explore would be testing popular consumer wrist sensors that claim to accurately detect emotion. Finally, a possible improvement to this study could have been conducting nested cross-validation, in which the hyperparameters would be window size/shift, sampling frequency, etc, so a wider range of views on the data are explored.

# Appendix

## Tables

```{r Data Dictionary, eval = TRUE}
collapse_rows_dt <- data.frame(Metric = c(rep("ECG (mV)", 3), rep("EDA ($\\mu$s)", 2), rep("EMG (mV)", 6), rep("Temp (C$^\\circ$)", 4),
                                      rep("Resp ($\\%$)", 3), rep("BVP", 4), rep("EDA ($\\mu$s)", 2), rep("Temp (C$^\\circ$)", 2)),
                               Description = linebreak(c(rep("Electrocardiography: 700 Hz", 3), 
                                                         rep("Electrodermal Activity: 700 Hz", 2), 
                                                         rep("Electromyography: 700 Hz", 6),
                                                         rep("Body Temperature: 700 Hz", 4), 
                                                         rep("Respiration: 700 Hz", 3), 
                                                         rep("Blood Volume Pulse: 64 Hz", 4), 
                                                         rep("Electrodermal Activity: 4 Hz", 2), 
                                                         rep("Skin Temperature: 4 Hz", 2))),
                               MSR = linebreak(c(rep("50 Hz", 3), 
                                                         rep("4 Hz", 2), 
                                                         rep("30 Hz", 6),
                                                         rep("1 Hz", 4), 
                                                         rep("10 Hz", 3), 
                                                         rep("1 Hz (HR)", 4), 
                                                         rep("4 Hz", 2), 
                                                         rep("1 Hz", 2))),
                               Features = linebreak(c("Mean", "Std Deviation", "Peaks", 
                                        "Mean" , "Std Deviation", 
                                        "Mean", "Std Deviation", "Range", "Median", "10th Percentile", "90th Percentile", 
                                        "Mean", "Std Deviation", "Range", "Median", 
                                        "Mean",  "Std Deviation", "Peaks", 
                                        "Mean", "Std Deviation", "Heart Rate Variability", "RMSSD of HRV",
                                        "Mean", "Std Deviation", 
                                        "Mean", "Std Deviation")))
kable(collapse_rows_dt, escape=F, booktabs = T, align = "llcc", caption = "Data Dictionary") %>%
  kable_styling(latex_options = c("hold_position")) %>% 
  group_rows(index = c("RespiBAN Chest Sensor" = 18, "Empatica E4 Wrist Sensor" = 8)) %>%
  column_spec(1, bold=T) %>%
  collapse_rows(columns = 1:3, latex_hline = "major", valign = "middle")
```

Note that while both sensors also included accelerometer data, it was not included in this analysis as it was only used to align the data points and distinguish the timestamps of the amusement and stress states. Other studies also have considered this variable and ultimately found it not to be useful, as you can read here [[9]][Bibliography].

For our final model, we decided to use linear interpolation to upsample the the electrocardiography *Empatica E4* data [[8]][Bibliography]. For example, say we observe the values $[2.0, 6.0, 12.0, 24.0]$ with 2 Hz as the current sampling rate (i.e. 2 and 6 are observations from the first second), and we wish to upsample to 4 Hz. We would translate these values to $[2.0, 3.33, 4.66, 6.0, 12.0, 16.0, 20.0, 24.0]$. We tested our model's sensitivity to this decison by also using forward projection (in the example above, forward projection would yield $[2.0, 2.0, 6.0, 6.0, 12.0, 12.0, 16.0, 16.0]$) and saw similar results, which is located in Table ** in the Appendix (REMOVE IF FALSE).

## Plots

#### Distributions of EDA and Temperature Measured at Wrist

```{r fig.width=10, fig.height=4, message=F, echo=F}
# distribution of average body temperature measured at wrist (window = 60s, shift = 0.25s)
ggplot(wesad) +
  geom_histogram(aes(x=TEMP_EMP4_MEAN, fill=label),alpha = 0.6, position = "identity")+
  facet_wrap(~subject, nrow = 3, scales = "free")+
  labs(x = "Body Temperature (??C)",
       title = "Empatica E4 Skin Temperature") 

# distribution of average electrodermal activity measured at wrist (window = 60s, shift = 0.25s)
ggplot(wesad) +
  geom_histogram(aes(x=EDA_EMP4_MEAN, fill=label),alpha = 0.6, position = "identity")+
  facet_wrap(~subject, nrow = 3, scales = "free")+
  labs(x = expression(paste("Electrodermal Activity (",mu,"S)")),
       title = "Empatica E4 Electrodermal Activity")
```

## Bibliography

1. Wray, S. (2017, July 27). Explosion of smart, wearable devices will triple revenue by 2016. Retrieved September 20, 2020, from https://inform.tmforum.org/features-and-analysis/2014/09/explosion-smart-wearable-devices-will-triple-revenue-2016/
2. Lu, T., Fu, C., Ma, M., Fang, C., & Turner, A. (2016, September 14). Healthcare Applications of Smart Watches. A Systematic Review. Retrieved September 20, 2020, from https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5052554/
3. Hiner, J. (2020, April 27). How Apple Watch helped fix my toxic level of stress. Retrieved September 20, 2020, from https://www.cnet.com/news/apple-watch-uncovered-my-toxic-stress-level-and-helped-me-fix-it/
4. 5 Things You Should Know About Stress. (n.d.). Retrieved September 20, 2020, from https://www.nimh.nih.gov/health/publications/stress/index.shtml
5. Clark, E. (2019, October 17). New Report Finds That More Than Half of Wearable Users Fear Inaccurate Health Data, Malfunctions; Users Cautioned Against Relying on Devices. Retrieved September 20, 2020, from https://www.prnewswire.com/news-releases/new-report-finds-that-more-than-half-of-wearable-users-fear-inaccurate-health-data-malfunctions-users-cautioned-against-relying-on-devices-300940397.html
6. Weintraub, K. (2016, October 13). Wearable health monitors not always reliable, study shows. Retrieved September 20, 2020, from https://www.usatoday.com/story/news/2016/10/12/wearable-health-monitors-not-always-reliable-study-shows/91922858/
7. Philip Schmidt, Attila Reiss, Robert Duerichen, Claus Marberger, and Kristof Van Laerhoven. 2018. Introducing WESAD, a Multimodal Dataset for Wearable Stress and Affect Detection. In Proceedings of the 20th ACM International Conference on Multimodal Interaction (ICMI '18). Association for Computing Machinery, New York, NY, USA, 400???408. DOI:https://doi.org/10.1145/3242969.3242985
8. https://www.cs.toronto.edu/~guerzhoy/320/lec/upsampling.pdf
9. Gjoreski, M., Lu??trek, M., Gams, M., &amp; Gjoreski, H. (2017, August 10). Monitoring stress with a wrist device using context. Retrieved September 27, 2020, from https://www.sciencedirect.com/science/article/pii/S1532046417301855
10. Mahdiani S, Jeyhani V, Peltokangas M, Vehkaoja A. Is 50 Hz high enough ECG sampling frequency for accurate HRV analysis? Annu Int Conf IEEE Eng Med Biol Soc. 2015;2015:5948-51. doi: 10.1109/EMBC.2015.7319746. PMID: 26737646.
11. Braithwaite, J. J., Watson, D. G., Jones, R., &amp; Rowe, M. (2013). A Guide for Analysing Electrodermal Activity (EDA) &amp; Skin Conductance Responses (SCRs) for Psychological Experiments. Retrieved September 26, 2020, from https://www.biopac.com/wp-content/uploads/EDA-SCR-Analysis.pdf
12. Pourmohammadi, S., &amp; Maleki, A. (2020, May 05). Stress detection using ECG and EMG signals: A comprehensive study. Retrieved September 27, 2020, from https://www.sciencedirect.com/science/article/pii/S0169260719320528
