---
title: "Case Study 2 - EDA"
author: "Alice Liao"
date: "9/19/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(cowplot)
```

# S2 Raw Data
```{r eval=F}
# load E4 ACC data, sample rate = 32Hz
# values are in 1/64 g
S2.E4.ACC <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/WESAD/S2/S2_E4_Data/ACC.csv",header = F)
S2.E4.ACC <- S2.E4.ACC %>%
    slice(-c(1,2)) %>%
  transmute(X = V1,
            Y = V2,
            Z = V3,
            time = 1:(nrow(S2.E4.ACC)-2))
  
head(S2.E4.ACC)
nrow(S2.E4.ACC)
```

```{r eval=F}
ggplot(S2.E4.ACC) +
  geom_line(aes(x = time, y = X/64), color = "blue") +
  labs(x = "Time", y = "X, unit in g")

ggplot(S2.E4.ACC) +
  geom_line(aes(x = time, y = Y/64))+
    labs(x = "Time", y = "Y, unit in g")

ggplot(S2.E4.ACC) +
  geom_line(aes(x = time, y = Z/64))+
    labs(x = "Time", y = "Z, unit in g")
```

```{r fig.width=10, fig.height=4,eval=F}
ggplot(S2.E4.ACC) +
  geom_line(aes(x = time, y = X/64), color = "dark blue") +
  geom_line(aes(x = time, y = Y/64), color = "red") + 
  geom_line(aes(x = time, y = Z/64), color = "dark green") +
  labs(y = "Accelerometer, unit in g",
       title = "Subject 2")
```

```{r eval=F}
# time format is minute.second
# ignore sRead, fRead
S2.Truth <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/WESAD/S2/S2_quest.csv", sep = ";", header = F)
head(S2.Truth)

S2.conditions <- S2.Truth %>%
  select(V1, V2, V3, V4, V5, V6) %>%
  slice(c(2:4))
S2.conditions <- as.data.frame(t(S2.conditions)) %>%
  slice(-1) %>%
  transmute(condition = V1,
            start = V2,
            end = V3)

```

```{r fig.height=5, fig.width=8,eval=F}
S2.E4.IBI <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/WESAD/S2/S2_E4_Data/IBI.csv",
                      header = F)
S2.E4.IBI <- S2.E4.IBI %>%
  slice(-1) %>%
  transmute(start = V1,
            IBI = as.numeric(V2))
p0.IBI.S2 = ggplot(data = S2.E4.IBI) +
  geom_histogram(aes(x = IBI)) +
  labs(title = "S2: Distribution of IBI")

S3.E4.IBI <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/WESAD/S3/S3_E4_Data/IBI.csv",
                      header = F)
S3.E4.IBI <- S3.E4.IBI %>%
  slice(-1) %>%
  transmute(start = V1,
            IBI = as.numeric(V2))
p0.IBI.S3 = ggplot(data = S3.E4.IBI) +
  geom_histogram(aes(x = IBI)) +
  labs(title = "S3: Distribution of IBI")

plot_grid(p0.IBI.S2,p0.IBI.S3, nrow = 2)
```


# EDA for merged Subject 2 and Subject 3 Dataset (Downsampled)

```{r}
# load merged csv (subject 2 & subject 3)
S2_S3_merged <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/merged_down_wesad.csv", header = T)
S2_S3_merged <- S2_S3_merged %>%
  mutate(subject = as.factor(subject),
         label = as.factor(label))

S2_merged <- S2_S3_merged %>%
  filter(subject == "S2") %>%
  select(-subject)
summary(S2_merged)

S3_merged <- S2_S3_merged %>%
  filter(subject == "S3") %>%
  select(-subject)
summary(S3_merged)
```


```{r S2-temperature, fig.height=5, fig.width=8}
p1 = ggplot(S2_merged) +
  geom_histogram(aes(x = Temp, fill = label)) +
  labs(title = "S2 Chest Measured Body Temperature")

p2 = ggplot(S2_merged) +
  geom_histogram(aes(x = TEMP_EMP4, fill = label)) +
  labs(title = "S2 Wrist Measured Body Temperature")

plot_grid(p1, p2, nrow = 2)
```


```{r S2-temperature-corr, fig.height=3, fig.width=5}
ggplot(S2_merged, aes(x = Temp, y = TEMP_EMP4)) +
  geom_point(aes(color=label)) +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Chest Body Temperature", y = "Wrist Body Temperature",
       title = "S2: Wrist vs. Chest Temprature")
```



```{r S3-temperature, fig.height=5, fig.width=8}

p3 = ggplot(S3_merged) +
  geom_histogram(aes(x = Temp, fill = label)) +
  labs(title = "S3 Chest Measured Body Temperature")

p4 = ggplot(S3_merged) +
  geom_histogram(aes(x = TEMP_EMP4, fill = label)) +
  labs(title = "S3 Wrist Measured Body Temperature")

plot_grid(p3, p4, nrow = 2)
```

```{r}
# Which rows from raw data have wierd temp data??
S2_S3_raw <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/merged_wesad.csv", header = T)
 
#83 rows, ~ 0.1 second chest sensors malfunctioned
which(S2_S3_raw$Temp <30)
```


```{r S3-temperature-corr, fig.height=3, fig.width=5}
ggplot(S3_merged, aes(x = Temp, y = TEMP_EMP4)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Chest Body Temperature", y = "Wrist Body Temperature",
       title = "S3: Wrist vs Chest Temperature") 
```
For both S2 and S3, wrist temperature data suggests clear distinction under amused and stressed conditions. 

```{r EMG, fig.height=4, fig.width=10}
p5 = ggplot(S2_merged) +
  geom_histogram(aes(x = EMG, fill = label)) +
  labs(title = "S2 Chest EMG")

p6 = ggplot(S3_merged) +
  geom_histogram(aes(x = EMG, fill = label)) +
  labs(title = "S3 Chest EMG")

plot_grid(p5,p6)
```
S3 EMG distribution when amused seems to have two modes; the distrubtions are very different for S2 and S3 --> heterogeity?

```{r ECG,fig.height=4, fig.width=10}
p7 = ggplot(S2_merged) +
  geom_histogram(aes(x = ECG, fill = label)) +
  labs(title = "S2 Chest ECG")

p8 = ggplot(S3_merged) +
  geom_histogram(aes(x = ECG, fill = label)) +
  labs(title = "S3 Chest ECG")

plot_grid(p7,p8)
```

```{r S2-EDA,fig.height=4, fig.width=10}
p9 = ggplot(S2_merged) +
  geom_histogram(aes(x = EDA, fill = label)) +
  labs(title = "S2 Chest EDA")

p10 = ggplot(S2_merged) +
  geom_histogram(aes(x = EDA_EMP4, fill = label)) +
  labs(title = "S2 Wrist EDA")

plot_grid(p9,p10)
```


```{r S2-EDA-corr,fig.height=3, fig.width=5}
ggplot(S2_merged, aes(x = EDA, y = EDA_EMP4)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Chest EDA", y = "Wrist EDA",
       title = "S2: Wrist vs Chest EDA") 
```


```{r S3-EDA,fig.height=4, fig.width=10}
p11 = ggplot(S3_merged) +
  geom_histogram(aes(x = EDA, fill = label)) +
  labs(title = "S3 Chest EDA")

p12 = ggplot(S3_merged) +
  geom_histogram(aes(x = EDA_EMP4, fill = label)) +
  labs(title = "S3 Wrist EDA")

plot_grid(p11,p12)
```
EDA distributions look different chest and wrist and for different subjects too. S3 in general has higher EDA; when stressed there is more sweat. 

```{r S3-EDA-corr,fig.height=3, fig.width=5}
ggplot(S3_merged, aes(x = EDA, y = EDA_EMP4)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Chest EDA", y = "Wrist EDA",
       title = "S3: Wrist vs Chest EDA") 
```


```{r resp,fig.height=5, fig.width=8}
p13 = ggplot(S2_merged) +
  geom_histogram(aes(x = Resp, fill = label)) +
  labs(title = "S2 Chest Respiration")

p14 = ggplot(S3_merged) +
  geom_histogram(aes(x = Resp, fill = label)) +
  labs(title = "S3 Chest Respiration")

plot_grid(p13,p14, nrow = 2)
```

```{r S2-HR-SRV,fig.height=5, fig.width=8}
p15 = ggplot(S3_merged) +
  geom_histogram(aes(x = HR_EMP4, fill = label)) +
  labs(title = "S3 Wrist Heart Rate")

p16 = ggplot(S3_merged) +
  geom_histogram(aes(x = HRV, fill = label)) +
  labs(title = "S3 Wrist Heart Rate Variability")

plot_grid(p15,p16, nrow = 2)
```


# PCA

```{r S2-PCA, echo=F}
S2_merged_PCA <- S2_merged %>%
  select(-c(X,label))

apply(S2_merged_PCA, 2, var)

S2_pr.out <- prcomp(S2_merged_PCA, scale = TRUE) # to standardize predictors

#S2_pr.out$rotation
S2_pr.var <- S2_pr.out$sdev^2 #the variance explained by each principal component
S2_pve <- S2_pr.var/sum(S2_pr.var)

plot(S2_pve, xlab = "Principal Component",
     ylab = "Proportion of Variance Explained", ylim = c(0,1), type = 'b',
     main = "Principal Component Analysis for Subject 2")
```

```{r S3-PCA, echo=F}
S3_merged_PCA <- S3_merged %>%
  select(-c(X,label))

apply(S3_merged_PCA, 2, var)

S3_pr.out <- prcomp(S3_merged_PCA, scale = TRUE) # to standardize predictors

#S3_pr.out$rotation
S3_pr.var <- S3_pr.out$sdev^2 #the variance explained by each principal component
S3_pve <- S3_pr.var/sum(S3_pr.var)

plot(S3_pve, xlab = "Principal Component",
     ylab = "Proportion of Variance Explained", ylim = c(0,1), type = 'b',
     main = "Principal Component Analysis for Subject 3")
```

# EDA Functions

```{r plot-functions}
# function to plot distributions of body temp measured at chest and wrist
get_temp_plots <- function(subject, data){
  p_temp_1 = ggplot(data) +
    geom_histogram(aes(x = Temp, fill = label)) +
    labs(title = paste0(subject," Chest Body Temperature"))
  
  p_temp_2 = ggplot(data) +
    geom_histogram(aes(x = TEMP_EMP4, fill = label)) +
    labs(title = paste0(subject," Wrist Body Temperature"))
  
  plot_grid(p_temp_1, p_temp_2, nrow = 2)
}

# function to plot wrist v.s. check temperature
get_temp_corr <- function(subject, data){
  ggplot(data, aes(x = Temp, y = TEMP_EMP4)) +
    geom_point(aes(color=label)) +
    geom_smooth(method = "lm", se = F) +
    labs(x = "Chest Body Temperature", y = "Wrist Body Temperature",
         title = paste0(subject," Wrist vs. Chest Temprature" ))
}

# function to plot distribution of EMG
get_EMG_plot <- function(subject, data){
  ggplot(data) +
    geom_histogram(aes(x = EMG, fill = label)) +
    labs(title = paste0(subject," Chest EMG"))
}

# function to plot distribution of ECG

get_ECG_plot <- function(subject, data){
  ggplot(data) +
    geom_histogram(aes(x = ECG, fill = label)) +
    labs(title = paste0(subject," Chest ECG"))
}

# function to plot distributions of EDA measured at chest and wrist
get_EDA_plots <- function(subject, data){
  p_EDA_1 = ggplot(data) +
    geom_histogram(aes(x = EDA, fill = label)) +
    labs(title = paste0(subject," Chest EDA"))
  
  p_EDA_2 = ggplot(data) +
    geom_histogram(aes(x = EDA_EMP4, fill = label)) +
    labs(title = paste0(subject," Wrist EDA"))
  
  plot_grid(p_EDA_1,p_EDA_2)
}

get_EDA_corr <- function(subject, data){
ggplot(data, aes(x = EDA, y = EDA_EMP4)) +
  geom_point(aes(color=label)) +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Chest EDA", y = "Wrist EDA",
       title = paste0(subject," Wrist vs Chest EDA"))
}

# function to plot distribution of respiration
get_resp_plot <- function(subject, data){
  ggplot(data) +
    geom_histogram(aes(x = Resp, fill = label)) +
    labs(title = paste0(subject," Chest Respiration"))
}

# function to plot the distribution of heart rate and HRV (SDNN method)
get_HR_plots <- function(subject, data){
  p_HR = ggplot(data) +
    geom_histogram(aes(x = HR_EMP4, fill = label)) +
    labs(title = paste0(subject," Wrist Heart Rate"))
  
  p_HR_time = ggplot(data) +
    geom_line(aes(x = X, y = HR_EMP4, color = label)) +
    labs(x = "Time",title = paste0(subject," in Hear Rate"))
  
  p_HRV = ggplot(data) +
  geom_histogram(aes(x = HRV, fill = label)) +
  labs(title = paste0(subject," Wrist Heart Rate Variability"))
  
  plot_grid(p_HR,p_HR_time,p_HRV, nrow = 3)
}


```

# EDA for merged Subject 4 (Downsampled) 
```{r load-data-S4,include=F}
S4_S5_S6_merged <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/merged_down_wesad_4_5_6.csv", header = T)

S4_S5_S6_merged <- S4_S5_S6_merged %>%
  mutate(subject = as.factor(subject),
         label = as.factor(label))
summary(S4_S5_S6_merged)

```

```{r S4, eval=F warning=F,message=F}
S4_merged <- S4_S5_S6_merged %>%
  filter(subject == "S4") %>% 
  select(-subject)
summary(S4_merged)

get_temp_plots("S4",S4_merged)
get_temp_corr("S4", S4_merged)
get_EMG_plot("S4",S4_merged)
get_ECG_plot("S4", S4_merged)
get_EDA_plots("S4", S4_merged)
get_EDA_corr("S4",S4_merged)
get_resp_plot("S4",S4_merged)
get_HR_plots("S4",S4_merged)
```


```{r}
S5_merged <- S4_S5_S6_merged %>%
  filter(subject == "S5") %>%
  select(-subject)
summary(S5_merged)

get_temp_plots("S5",S5_merged)
get_temp_corr("S5", S5_merged)
get_EMG_plot("S5",S5_merged)
get_ECG_plot("S5", S5_merged)
get_EDA_plots("S5", S5_merged)
get_EDA_corr("S5",S5_merged)
get_resp_plot("S5",S5_merged)
get_HR_plots("S5",S5_merged)
```
# EDA for Subject 6
```{r, message=F, warning=F, echo=F}
S6_merged <- S4_S5_S6_merged %>%
  filter(subject == "S6") %>%
  select(-subject)
#summary(S6_merged)

get_temp_plots("S6",S6_merged)
get_temp_corr("S6", S6_merged)
get_EMG_plot("S6",S6_merged)
get_ECG_plot("S6", S6_merged)
get_EDA_plots("S6", S6_merged)
get_EDA_corr("S6",S6_merged)
get_resp_plot("S6",S6_merged)
get_HR_plots("S6",S6_merged)
```

```{r load-data-S7-S8-S9,include=F}
S7_S8_S9_merged <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/merged_down_wesad_7_8_9.csv", header = T)

S7_S8_S9_merged <- S7_S8_S9_merged %>%
  mutate(subject = as.factor(subject),
         label = as.factor(label))
summary(S7_S8_S9_merged)
```

# EDA for Subject 7 
```{r S7, message=F, warning=F,echo=F}
# need to amend this
S7_merged <- S7_S8_S9_merged %>%
  filter(subject == "S7") %>% 
  select(-subject)

summary(S7_merged)

get_temp_plots("S7",S7_merged)
get_temp_corr("S7", S7_merged)
get_EMG_plot("S7",S7_merged)
get_ECG_plot("S7", S7_merged)
get_EDA_plots("S7", S7_merged)
get_EDA_corr("S7",S7_merged)
get_resp_plot("S7",S7_merged)
get_HR_plots("S7",S7_merged)
```

# EDA for Subject 8

#### NOTE: Subject had already a rather stressful day, prior to the study.Subject felt rather cold in the room where the stress condition was carried out.
```{r S8, warning=F, message=F, echo=F}
S8_merged <- S7_S8_S9_merged %>%
  filter(subject == "S8") %>% 
  select(-subject)
summary(S8_merged)

get_temp_plots("S8",S8_merged)
get_temp_corr("S8", S8_merged)
get_EMG_plot("S8",S8_merged)
get_ECG_plot("S8", S8_merged)
get_EDA_plots("S8", S8_merged)
get_EDA_corr("S8",S8_merged)
get_resp_plot("S8",S8_merged)
get_HR_plots("S8",S8_merged)
```
# EDA for Subject 9

```{r S9, warning=F, message=F,echo=F}
S9_merged <- S7_S8_S9_merged %>%
  filter(subject == "S9") %>% 
  select(-subject)
summary(S9_merged)

get_temp_plots("S9",S9_merged)
get_temp_corr("S9", S9_merged)
get_EMG_plot("S9",S9_merged)
get_ECG_plot("S9", S9_merged)
get_EDA_plots("S9", S9_merged)
get_EDA_corr("S9",S9_merged)
get_resp_plot("S9",S9_merged)
get_HR_plots("S9",S9_merged)
```

```{r load-data-S10-S11-S13,include=F}
S10_S11_S13_merged <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/merged_down_wesad_10_11_13.csv", header = T)

S10_S11_S13_merged <- S10_S11_S13_merged %>%
  mutate(subject = as.factor(subject),
         label = as.factor(label))

summary(S10_S11_S13_merged)
```
# EDA for Subject 10 

```{r S10, warning=F, message=F,echo=F}
S10_merged <- S10_S11_S13_merged %>%
  filter(subject == "S10") %>% 
  select(-subject)
summary(S10_merged)

get_temp_plots("S10",S10_merged)
get_temp_corr("S10", S10_merged)
get_EMG_plot("S10",S10_merged)
get_ECG_plot("S10", S10_merged)
get_EDA_plots("S10", S10_merged)
get_EDA_corr("S10",S10_merged)
get_resp_plot("S10",S10_merged)
get_HR_plots("S10",S10_merged)
```

# EDA for Subject 11

```{r S11, warning=F, message=F,echo=F}
S11_merged <- S10_S11_S13_merged %>%
  filter(subject == "S11") %>% 
  select(-subject)
summary(S11_merged)

get_temp_plots("S11",S11_merged)
get_temp_corr("S11", S11_merged)
get_EMG_plot("S11",S11_merged)
get_ECG_plot("S11", S11_merged)
get_EDA_plots("S11", S11_merged)
get_EDA_corr("S11",S11_merged)
get_resp_plot("S11",S11_merged)
get_HR_plots("S11",S11_merged)
```

# EDA for Subject 13

```{r S13, warning=F, message=F,echo=F}
S13_merged <- S10_S11_S13_merged %>%
  filter(subject == "S13") %>% 
  select(-subject)
summary(S13_merged)

get_temp_plots("S13",S13_merged)
get_temp_corr("S13", S13_merged)
get_EMG_plot("S13",S13_merged)
get_ECG_plot("S13", S13_merged)
get_EDA_plots("S13", S13_merged)
get_EDA_corr("S13",S13_merged)
get_resp_plot("S13",S13_merged)
get_HR_plots("S13",S13_merged)
```


```{r load-data-S14-S15-S16-S17,include=F}
S14_S15_S16_S17_merged <- read.csv("~/Desktop/2020 Fall/STA440/Case 2/merged_down_wesad_14_15_16_17.csv", 
                                   header = T)

S14_S15_S16_S17_merged <- S14_S15_S16_S17_merged %>%
  mutate(subject = as.factor(subject),
         label = as.factor(label))

summary(S14_S15_S16_S17_merged)
```

# EDA for Subject 14

```{r S14, warning=F, message=F,echo=F}
S14_merged <- S14_S15_S16_S17_merged %>%
  filter(subject == "S14") %>% 
  select(-subject)
summary(S14_merged)

get_temp_plots("S14",S14_merged)
get_temp_corr("S14", S14_merged)
get_EMG_plot("S14",S14_merged)
get_ECG_plot("S14", S14_merged)
get_EDA_plots("S14", S14_merged)
get_EDA_corr("S14",S14_merged)
get_resp_plot("S14",S14_merged)
get_HR_plots("S14",S14_merged)
```

# EDA for Subject 15

```{r S15, warning=F, message=F, echo=F}
S15_merged <- S14_S15_S16_S17_merged %>%
  filter(subject == "S15") %>% 
  select(-subject)
summary(S15_merged)

get_temp_plots("S15",S15_merged)
get_temp_corr("S15", S15_merged)
get_EMG_plot("S15",S15_merged)
get_ECG_plot("S15", S15_merged)
get_EDA_plots("S15", S15_merged)
get_EDA_corr("S15",S15_merged)
get_resp_plot("S15",S15_merged)
get_HR_plots("S15",S15_merged)
```

# EDA for Subject 16

```{r S16, warning=F, message=F, echo=F}
S16_merged <- S14_S15_S16_S17_merged %>%
  filter(subject == "S16") %>% 
  select(-subject)
summary(S16_merged)

get_temp_plots("S16",S16_merged)
get_temp_corr("S16", S16_merged)
get_EMG_plot("S16",S16_merged)
get_ECG_plot("S16", S16_merged)
get_EDA_plots("S16", S16_merged)
get_EDA_corr("S16",S16_merged)
get_resp_plot("S16",S16_merged)
get_HR_plots("S16",S16_merged)
```

# EDA for Subject 17

```{r S16, warning=F, message=F,echo=F}
S17_merged <- S14_S15_S16_S17_merged %>%
  filter(subject == "S17") %>% 
  select(-subject)
summary(S17_merged)

get_temp_plots("S17",S17_merged)
get_temp_corr("S17", S17_merged)
get_EMG_plot("S17",S17_merged)
get_ECG_plot("S17", S17_merged)
get_EDA_plots("S17", S17_merged)
get_EDA_corr("S17",S17_merged)
get_resp_plot("S17",S17_merged)
get_HR_plots("S17",S17_merged)
```

